The human in the loop: volunteer-based metacomputers as a socio-technical system

### 2016-04-23 Alife16/paper.tex
minor issues
References #116

\begin{abstract}
Volunteer computing is a form of distributed computing where users
decide on their participation and the amount of time and other
resources they will ``lend''. This makes them an essential part of the
algorithm and of the performance of the whole system. As a
sociotechnical system, this participation follows some patterns and in
this paper we examine the result of several volunteer distributed
evolutionary computation experiments and try to find out which are
those patterns and what makes an experiment successful or not,
including the feedback loop that is created between the users and the
algorithm itself. In this paper we will examine what kind of patterns
we can expect and what influence those patterns have on the
implementation of the evolutionary algorithm.
\end{abstract}

@@

On the other hand, we are also interested in measuring the performance
of volunteer computing systems, an area in which there have been
relatively few efforts.
There were some initial attempts to avoid the differences in performance
that could be obtained from volunteers  by making
the algorithm adaptive to the kind of resources allotted to it
\citep{milani2004online}, which is actually not such a big problem in
algorithms such as the EA that can be easily 
parallelized via population splitting or by farming out the evaluation to all
the nodes available. Lately, several approaches have focused on the
fault-tolerance of volunteer systems
\citep{gonzalez2010characterizing} which it can, of course, be studied in
the more general context of distributed computing 
\citep{nogueras2015studying} or including it in a more general study of the
performance of the EA itself
\citep{DBLP:journals/gpem/LaredoBGVAGF14}. This performance cannot be
measured without first understanding the dynamics of this kind of systems. Initial
work was done for peer to peer systems by Stutzbach et
al. \citep{stutzbach2006understanding} and extended to volunteer
computing by Laredo et al. \citep{churn08,laredo2008rcp}. A similar
study was performed by Martínez et al. on the Capataz system
\citep{martinez2015capataz}; however, in this case the number of
computers used was known in advance and the main focus was on
measuring the speed up and how job bundling helped to reduce overhead and
enhance performance. On the contrary, in this paper, we will use {\em
  actual} volunteers, anonymous persons that are requested to join an
experiment through social networks.

@@

In general, a distributed volunteer-based evolutionary computation
system based on the browser is simply a client-server system
whose client is, or it can be, embedded in the browser via
JavaScript. Since JavaScript is  the only language that is present
across all browsers, the choice was quite clear. We should emphasize
that NodIO is more intended as an auxiliary computing engine, more
than the main one, so performance of JavaScript as a language is not
so important; even so, we have made a comparison between JavaScript and other languages
\citep{2015arXiv151101088M} that shows that performance of JavaScript
is comparable to other interpreted languages; compiled languages would
be faster, but, of course, it is impossible to gather volunteers
spontaneously and without any installation with them.

@@

Thus, {\sf NodIO} architecture has two tiers:\begin{enumerate}
\item A REST server, that is, a server that includes several {\em
  routes} 
  that can be called for storing and retrieving information (the ''CRUD'' cycle:
  create, request, update, and delete) from the server. 
  A JSON data format is used for the communication between 
  clients and the server. There are two kinds of information:
  {\em problem} based, that is, related to the
  evolutionary algorithm such as {\tt PUT}ing a chromosome in or {\tt
  GET}ing a random chromosome from it, and {\em information} related
  to the performance and state of the experiments. It also performs logging
  duties, but they are basically a very lightweight and high performance
  data storage \citep{jj:idc:lowcost}.
  The server has the capability to
  run a single experiment, storing the chromosomes in a data structure
  that is reset when the solution is found. This data structure can
  hold every chromosome in a particular experiment, or have a finite
  size that erases the oldest chromosomes once it has filled to
  capacity. In this paper we will test both implementations.  
\item A client that includes the evolutionary algorithm as
  JavaScript code embedded in a web page that displays graphs, some
  additional links, and information on the experiment. This code runs
  an evolutionary algorithm {\em island} that starts with a random
  population, then after every 100 generations, it sends the best individual
  back to the server (via a {\tt PUT} request), and then requests a random
  individual back from the server (via a {\tt GET} request). We have
  kept the number of generations between migrations fixed since it is
  a way of finding out how many real work every client has done. 
\end{enumerate}

Figure \ref{fig:system} describes the general system architecture and
algorithm behavior. Different web technologies, such as JQuery or {\tt
  Chart.js} have been used to build the user interface elements of the
framework, a part of which is shown in Figure \ref{fig:screenshot} and
should be running in \url{http://nodio-jmerelo.rhcloud.com}.

@@

\begin{table}[htb]
\caption{Experiment table, with summary of results. \label{tab:runs}}
\begin{center}
\begin{tabular}{l|rrl}
\hline
Experiment & \#Runs & Different IPs & Traps \\
\hline
April 4th 4/4 & 57 & 191 & 40 \\
April 24th 4/24 &  231 & 559& 40  \\
July 31th 7/31 & 97 & 179 & 40 \\
\hline
February, cache=128 & 61 & 75 & 50  \\
February, cache=64 & 61 & 220 & 50  \\
February, cache=32 & 39 & 86 & 50  \\
% grep start ~/Code/splash-volunteer/data/2016/nodio-2016-2-28-cache=32.log
\hline
\end{tabular}
\end{center}
\end{table}
%
Initial experiments were set up using the OpenShift
PaaS, which provides a free tier, making the whole experiment cost
equal to \$0.00. Experiments were
announced through a series of posts on Twitter and, in the latest case, Telegram, and
results were published here \citep{DBLP:conf/gecco/GuervosG15}. For the
purpose of this paper, we repeated the announcement several times
through the month of February. All  %% April --> January ??
in all, we have the set of runs with the characteristics shown in
Table \ref{tab:runs}. In general, every experiment took several
days. No particular care was taken about the time of the announcement
%Pablo: but it would great at least describe how many times was announced and how long
or the particular wording. Every {\em experiment} consisted in running
until the solution of the 50-trap problem was found. When the correct
solution was sent to the server, the counter was updated and the pool
of solutions resets to the void set. There was no special intention to wait
until all clients had finished, thus it might happen that. In fact,
the islands running in the browser {\em spill} from one experiment to
the next. However, previous experiments have proven that the influence
of these islands in the next experiment is indeed negligible. %Pablo: cite

@@

The number of {\tt PUTs}, every one corresponding to 100 generations,
is the algorithmic result. It is relatively unchanged for the first
batch and around 20, that is, 2000 generations or 2000*128 = 256000
evaluations. In this case, an ``unlimited'' cache was used, with all
individuals sent from clients stored until the end of the
experiment. However, we were interested in measuring also the
performance of the algorithm itself by changing the cache size, after
making it limited. As it can be seen in the table, there is a clear
%Pablo: which table?
change in the number of evaluations needed, with smaller cache sizes
producing solutions in less evaluations, until it is for the cache
size = 32 roughly twice as much as with the previous problem, with 40
traps. This is a good result and is also algorithmically consistent
with other results obtained using the same type of problems. Since in
this paper we were interested in leveraging the user's CPU cycles by
improving the algorithm, a good conclusion of this paper is that
having a small pool size helps clients to obtain ``good'' individuals
from the pool, as opposed to any individual that could be obtained
before. Besides, cache policy deletes the oldest individuals, which
makes those in the pool be {\em current}, helping then newcomers and %Pablo: "then helping newcomers", instead?
any participant obtain the best individuals found in the last part of
the experiment. Besides, a limited cache helps also in cases with a
bigger search space or longer running times when the server simply
crashed due to lack of RAM.

\begin{figure}[!htb]
\centering
\includegraphics[width=0.32\linewidth]{time-vs-ips-OS-4-4.png}
\includegraphics[width=0.32\linewidth]{time-vs-ips-OS-4-24.png}
\includegraphics[width=0.32\linewidth]{time-vs-ips-OS-7-31.png}
\includegraphics[width=0.32\linewidth]{time-vs-ips-alife-128.png}
\includegraphics[width=0.32\linewidth]{time-vs-ips-alife-64.png}
\includegraphics[width=0.32\linewidth]{time-vs-ips-alife-32.png}
\caption{Duration of experiments vs. number of different IPs (nodes)
  participating in it, with averages and standard deviation shown as
  red dots; in the case there is a single red dot, there was a single
  experiment in which many computers participated (for instance, 16
  computers in the experiment in the far left or 29 in the middle
  one). 
Shades of blue indicate how many experiments included that many unique IPs,
so lighter shade for a column of dots indicates that a particular number
of computers happened less frequently, while darker shadow means more frequency. 
From left to right and top to bottom experiments 4/4, 4/24 and 7/31,
followed by experiments with 50 traps, cache=128, 64, 32.}
\label{fig:duration}
\end{figure}
%
We will have to analyze experimental data a bit further to find out why
this happens and also if there are some patterns in the three sets of
experiments. An interesting question to ask, for instance, is if
by adding more computers makes the experiment take less. In fact, as
shown in Figure \ref{fig:duration}, the {\em addition} of more computers does
not seem to contribute to decreasing the time needed to finish the
experiment. However, the cause-effect relationship is not clear at
all. It might be the opposite: since experiments take longer to finish
and might in fact be abandoned with no one contributing for some time,
the probability of someone new joining them is higher. In fact,
with experiments taking a few seconds and due to the way the
experiments are announced, it is quite difficult that several
volunteers join in in such a short period of time, even more if we take
into account that volunteers are not {\em carried over} from previous
experiments. 
%
\begin{figure}[!htb]
\centering
\includegraphics[width=0.95\linewidth]{ips-per-minute-cache=64.png}
\caption{Simultaneous IPs every minute of the experiment with cache =
  64. \label{fig:otisdriftwood}}
\end{figure}

That is why we used a more difficult problem in the second batch of
experiments, which is shown in the bottom row of Figure
\ref{fig:duration}. The pattern is remarkably similar, showing a
positive correlation between the time for solving the problem and the
number of computers, at least for cache sizes 128 and 64. However, it
is interesting to observe that, for cache=32, the time needed to find
the solution decreases from one to approximately 4-5 nodes, to then
increase for a higher number of participating computers, distinguished
by IP. The green dot at the bottom is probably an outlier that we
will try to explain later on. This leads us to conclude that a better
amount of computers might contribute to speed up the solution, if the
time the experiment ideally takes is sufficient, that is, of the order of a
minute, and enough volunteers concur simultaneously. This is also
observed, not so clearly, in the case of cache=64, with an interval of
around 10 IPs obtaining less time than experiments with less or more
IPs, and of the same order, between 10 and 100 seconds. If we look at
the graph that shows the number of IPs or volunteers per minute for
this experiment, shown in Figure \ref{fig:otisdriftwood}, we see that
there are peaks of more than 25 volunteers, and a period of several
hours with a minimum of 6 computers and peaks of more than 10. The
long period after midnight where there is a single volunteer left
masks the success achieved during this set of experiments, from which
we draw two lessons: first, you need a social network influencer to
%Pablo: people know the meaning of "influencer"? Probably yes, but I mention it just in case.
announce your experiments and second, no matter what, do not do any
experiment after midnight. This statement, which might seem tongue in
cheek, in fact, it is a conclusion drawn from the experimental data and
to what extent the social network is an essential part of the
description and performance of the NodIO volunteer computing system.

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=0.32\linewidth]{weibull-puts-openshift-4-4.png}
\includegraphics[width=0.32\linewidth]{weibull-puts-openshift-4-24.png}
\includegraphics[width=0.32\linewidth]{weibull-puts-openshift-7-31.png}
\includegraphics[width=0.32\linewidth]{weibull-fit-cache=128.png}
\includegraphics[width=0.32\linewidth]{weibull-fit-cache=64.png}
\includegraphics[width=0.32\linewidth]{weibull-fit-cache=32.png}
\caption{Number of {\tt PUT}s per unique IP and fit to a Weibull
  distribution (in red). From left to right and top to bottom, experiments
  4/4, 4/24 and 7/31 and new experiments with cache=128, 64, 32.} 
% Plotted with ../data/plot-zipf-openshift.R
\label{fig:puts:os}
\end{figure}
%
A similar exponential distribution also appears if we rank HTTP {\tt
  PUT}s, equivalents to the number of 
generations divided by 100, or to evaluations divided by 12800,
contributed by every user, which is shown 
in Figure \ref{fig:puts:os}. These results show a Zipf-like behavior,
that is, a power law with a small {\em bump} in the lowest
values. After testing the Generalized Extreme Value distribution and
failing for the new batch of experiments, we have fitted it to a
Weibull distribution \cite{thoman1969inferences} with the resulting
parameters shown in Table \ref{tab:puts:os}. 
%
\begin{table}
\caption{Weibull distribution  parameters of the fit of
  the number of {\tt PUT}s per unique IP. \label{tab:puts:os}}
\begin{center}
\begin{tabular}{l|cc}
\hline
Experiment  &  Scale $\sigma$ & Shape $\xi$ \\
\hline
4/4 &  43.07 $\pm$ 5.80 &  0.57 $\pm$ 0.03 \\
4/24 & 22.97 $\pm$ 1.57 & 0.66 $\pm$  0.02  \\
7/31 &  53.18 $\pm$ 7.77 &  0.54 $\pm$ 0.03   \\
\hline
Cache 128 & 205.28 $\pm$ 32.10 & 0.77 $\pm$ 0.07 \\ 
Cache 64 & 178.99 $\pm$ 36.44 & 0.60 $\pm$ 0.05 \\ 
Cache 32 & 168.15 $\pm$ 49.28 & 0.57 $\pm$ 0.07 \\
\hline
% from plot-zipf-openshift.R and ips-puts-alife.R
\end{tabular}
\end{center}
\end{table}
%
The inverse
Weibull distribution is a special case of the GEV distribution we have
used in those papers, and appears usually in natural sciences and
artificial life, usually related to decay. It has been frequently
fitted to volunteer computing frameworks such as SETI@home
\citep{javadi2009mining}. The model that user behavior follows can be
explained straighforwardly: when users visit the page, it draws their
attention for a limited amount of time. They give it a chance for a
few seconds. If something there amuses them or they can engage in a
conversation about it, they stay for a while longer, otherwise, they
leave. The {\em scale} parameter, which is around 20-40 in the first
batch for the 40 Trap problem and between 160 and 200 in the 60 traps
problem, depends mainly on the maximum number of generations people
leave it running. Since it finishes or stalls after a number of
generations shown in Figure \cite{tab:summary:os}, volunteers just
leave after that. Curiously enough, the scale parameter is roughly
twice the median number of {\tt PUT}s per experiment, showing that, on
average, the most loyal users reload the page twice after finishing or
after seeing the evolution does not progress. This rule of thumb
breaks down with the last experiment, however, which is interesting by
itself, too.

>>>>

\begin{abstract}
Volunteer computing is a form of distributed computing where users
decide on their participation and the amount of time and other
resources they will ``lend''. This makes them an essential part of the
algorithm and of the performance of the whole system. As a
sociotechnical system, this participation follows some patterns and in
this paper we examine the result of several volunteer distributed
evolutionary computation experiments and try to find out what
those patterns are and what makes an experiment successful or not,
including the feedback loop that is created between the users and the
algorithm itself.
\end{abstract}

@@

On the other hand, we are also interested in measuring the performance
of volunteer computing systems, an area in which there have been
relatively few efforts.
There were some initial attempts to avoid the differences in performance
that could be obtained from volunteers  by making
the algorithm adaptive to the kind of resources allotted to it
\citep{milani2004online}, which is actually not such a big problem in
algorithms such as the EA that can be easily 
parallelized via population splitting or by farming out the evaluation to all
the nodes available. Lately, several approaches have focused on the
fault-tolerance of volunteer systems
\citep{gonzalez2010characterizing} which it can, of course, be studied in
the more general context of distributed computing 
\citep{nogueras2015studying}  or included in a more general 
study of the performance of the EA itself
\citep{DBLP:journals/gpem/LaredoBGVAGF14}. This performance cannot be
measured without first understanding the dynamics of this kinds of systems. Initial
work was done for peer to peer systems by Stutzbach et
al. \citep{stutzbach2006understanding} and extended to volunteer
computing by Laredo et al. \citep{churn08,laredo2008rcp}. A similar
study was performed by Martínez et al. on the Capataz system
\citep{martinez2015capataz}; however, in this case the number of
computers used was known in advance and the main focus was on
measuring the speed up and how job bundling helped to reduce overhead and
enhance performance. On the contrary, in this paper, we will use {\em
  actual} volunteers, anonymous persons that are requested to join an
experiment through social networks.

@@

In general, a distributed volunteer-based evolutionary computation
system based on the browser is simply a client-server system
whose client is, or it can be, embedded in the browser via
JavaScript. Since JavaScript is  the only language that is present
across all browsers, the choice was quite clear. We should emphasize
that NodIO is more intended as an auxiliary computing engine, more
than the main one, so performance of JavaScript as a language is not
so important; even so, we have made a comparison between JavaScript and other languages
\citep{2015arXiv151101088M} that shows that the performance of JavaScript
is comparable to other interpreted languages; compiled languages would
be faster, but, of course, it is impossible to gather volunteers
spontaneously and without any installation with them.

@@

Thus, {\sf NodIO} architecture has two tiers:\begin{enumerate}
\item A REST server, that is, a server that includes several {\em
  routes} 
  that can be called for storing and retrieving information (the `CRUD'' cycle:
  create, request, update, and delete) from the server. 
  A JSON data format is used for the communication between 
  clients and the server. There are two kinds of information:
  {\em problem} based, that is, related to the
  evolutionary algorithm such as {\tt PUT}ing a chromosome in or {\tt
  GET}ing a random chromosome from it, and {\em information} related
  to the performance and state of the experiments. It also performs logging
  duties, but they are basically a very lightweight and high performance
  data storage \citep{jj:idc:lowcost}.
  The server has the capability to
  run a single experiment, storing the chromosomes in a data structure
  that is reset when the solution is found. This data structure can
  hold every chromosome in a particular experiment, or have a finite
  size that erases the oldest chromosomes once it has filled to
  capacity. In this paper we will test both implementations.  
\item A client that includes the evolutionary algorithm as
  JavaScript code embedded in a web page that displays graphs, some
  additional links, and information on the experiment. This code runs
  an evolutionary algorithm {\em island} that starts with a random
  population, then after every 100 generations, it sends the best individual
  back to the server (via a {\tt PUT} request), and then requests a random
  individual back from the server (via a {\tt GET} request). We have
  kept the number of generations between migrations fixed since it is
  a way of finding out how much real work every client has done. 
\end{enumerate}

Figure \ref{fig:system} describes the general system architecture and
algorithm behavior. Different web technologies, such as JQuery and {\tt
  Chart.js} have been used to build the user interface elements of the
framework, a part of which is shown in Figure \ref{fig:screenshot} and
should be running in \url{http://nodio-jmerelo.rhcloud.com}.

@@

\begin{table}[htb]
\caption{Experiment table, with summary of results. \label{tab:runs}}
\begin{center}
\begin{tabular}{l|rrl}
\hline
Experiment & \#Runs & Different IPs & Traps \\
\hline
April 4th 4/4 & 57 & 191 & 40 \\
April 24th 4/24 &  231 & 559& 40  \\
July 31th 7/31 & 97 & 179 & 40 \\
\hline
February, cache=128 & 61 & 75 & 50  \\
February, cache=64 & 61 & 220 & 50  \\
February, cache=32 & 39 & 86 & 50  \\
% grep start ~/Code/splash-volunteer/data/2016/nodio-2016-2-28-cache=32.log
\hline
\end{tabular}
\end{center}
\end{table}
%
Initial experiments were set up using the OpenShift
PaaS \url{www.openshift.com}, which provides a free tier, making the whole experiment cost
equal to \$0.00. Experiments were
announced through a series of posts on Twitter and, in the latest case, Telegram, and
results were published in\citep{DBLP:conf/gecco/GuervosG15}. For the
purpose of this paper, we repeated the announcement several times
through the month of February. All  %% April --> January ??
in all, we have the set of runs with the characteristics shown in
Table \ref{tab:runs}. In general, every experiment took several
days. No particular care was taken about the time of the announcement
%Pablo: but it would great at least describe how many times was announced and how long
or the particular wording. Every {\em experiment} consisted in running
until the solution of the 50-trap problem was found. When the correct
solution was sent to the server, the counter was updated and the pool
of solutions resets to the void set. There was no special intention to wait
until all clients had finished, thus it might happen that. In fact,
the islands running in the browser {\em spill} from one experiment to
the next. However, previous experiments have proven that the influence
of these islands in the next experiment is indeed negligible. %Pablo: cite

@@

The number of {\tt PUTs}, every one corresponding to 100 generations,
is the algorithmic result. It is relatively unchanged for the first
batch and around 20, that is, 2000 generations or 2000*128 = 256000
evaluations. In this case, an ``unlimited'' cache was used, with all
individuals sent from clients stored until the end of the
experiment. However, we were interested in measuring also the
performance of the algorithm itself by changing the cache size, after
making it limited. As it can be seen in the table, there is a clear
%Pablo: which table?
change in the number of evaluations needed, with smaller cache sizes
producing solutions in less evaluations, until it is for the cache
size = 32 roughly twice as much as with the previous problem, with 40
traps. This is a good result and is also algorithmically consistent
with other results obtained using the same type of problems. Since in
this paper we were interested in leveraging the user's CPU cycles by
improving the algorithm, a good conclusion of this paper is that
having a small pool size helps clients to obtain ``good'' individuals
from the pool, as opposed to any individual that could be obtained
before. Besides, the cache policy deletes the oldest individuals, which
makes those in the pool be {\em current}, helping then newcomers and %Pablo: "then helping newcomers", instead?
any participant obtain the best individuals found in the last part of
the experiment. Besides, a limited cache helps also in cases with a
bigger search space or longer running times when the server simply
crashed due to lack of RAM.

\begin{figure}[!htb]
\centering
\includegraphics[width=0.32\linewidth]{time-vs-ips-OS-4-4.png}
\includegraphics[width=0.32\linewidth]{time-vs-ips-OS-4-24.png}
\includegraphics[width=0.32\linewidth]{time-vs-ips-OS-7-31.png}
\includegraphics[width=0.32\linewidth]{time-vs-ips-alife-128.png}
\includegraphics[width=0.32\linewidth]{time-vs-ips-alife-64.png}
\includegraphics[width=0.32\linewidth]{time-vs-ips-alife-32.png}
\caption{Duration of experiments vs. number of different IPs (nodes)
  participating in it, with averages and standard deviation shown as
  red dots; in the case there is a single red dot, there was a single
  experiment in which that many computers participated (for instance, 16
  computers in the experiment in the far left or 29 in the middle
  one). 
Shades of blue indicate how many experiments included that many unique IPs,
so lighter shade for a column of dots indicates that a particular number
of computers happened less frequently, while darker shadow means more frequency. 
From left to right and top to bottom experiments 4/4, 4/24 and 7/31,
followed by experiments with 50 traps, cache=128, 64, 32.}
\label{fig:duration}
\end{figure}
%
We will have to analyze experimental data a bit further to find out why
this happens and also if there are some patterns in the three sets of
experiments. An interesting question to ask, for instance, is if 
by adding more computers it makes the experiment take less time. In fact, as
shown in Figure \ref{fig:duration}, the {\em addition} of more computers does
not seem to contribute to decreasing the time needed to finish the
experiment. However, the cause-effect relationship is not clear at
all. It might be the opposite: since experiments take longer to finish
and might in fact be abandoned with no one contributing for some time,
the probability of someone new joining them is higher. In fact,
with experiments taking a few seconds and due to the way the
experiments are announced, it is quite difficult that several
volunteers join in in such a short period of time, even more if we take
into account that volunteers are not {\em carried over} from previous
experiments. 
%
\begin{figure}[!htb]
\centering
\includegraphics[width=0.95\linewidth]{ips-per-minute-cache=64.png}
\caption{Simultaneous IPs every minute of the experiment with cache =
  64. \label{fig:otisdriftwood}}
\end{figure}

That is why we used a more difficult problem in the second batch of
experiments, which is shown in the bottom row of Figure
\ref{fig:duration}. The pattern is remarkably similar, showing a
positive correlation between the time for solving the problem and the
number of computers, at least for cache sizes 128 and 64. However, it
is interesting to observe that, for cache=32, the time needed to find
the solution decreases from one to approximately 4-5 nodes, to then
increase for a higher number of participating computers, distinguished
by IP. The green dot at the bottom is probably an outlier that we
will try to explain later on. This leads us to conclude that a larger
amount of computers might contribute to speed up the solution, if the
time the experiment ideally takes is sufficient, that is, of the order of a
minute, and enough volunteers concur simultaneously. This is also
observed, not so clearly, in the case of cache=64, with an interval of
around 10 IPs obtaining less time than experiments with less or more
IPs, and of the same order, between 10 and 100 seconds. If we look at
the graph that shows the number of IPs or volunteers per minute for
this experiment, shown in Figure \ref{fig:otisdriftwood}, we see that
there are peaks of more than 25 volunteers, and a period of several
hours with a minimum of 6 computers and peaks of more than 10. The
long period after midnight where there is a single volunteer left
masks the success achieved during this set of experiments, from which
we draw two lessons: first, you need a social network influencer to
%Pablo: people know the meaning of "influencer"? Probably yes, but I mention it just in case.
announce your experiments and second, no matter what, do not do any
experiment after midnight. This statement, which might seem tongue in
cheek, in fact, it is a conclusion drawn from the experimental data and
to what extent the social network is an essential part of the
description and performance of the NodIO volunteer computing system.

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=0.32\linewidth]{weibull-puts-openshift-4-4.png}
\includegraphics[width=0.32\linewidth]{weibull-puts-openshift-4-24.png}
\includegraphics[width=0.32\linewidth]{weibull-puts-openshift-7-31.png}
\includegraphics[width=0.32\linewidth]{weibull-fit-cache=128.png}
\includegraphics[width=0.32\linewidth]{weibull-fit-cache=64.png}
\includegraphics[width=0.32\linewidth]{weibull-fit-cache=32.png}
\caption{Number of {\tt PUT}s per unique IP and fit to a Weibull
  distribution (in red). From left to right and top to bottom, experiments
  4/4, 4/24 and 7/31 and new experiments with cache=128, 64, 32.} 
% Plotted with ../data/plot-zipf-openshift.R
\label{fig:puts:os}
\end{figure}
%
A similar exponential distribution also appears if we rank HTTP {\tt
  PUT}s, equivalents to the number of 
generations divided by 100, or to evaluations divided by 12800,
contributed by every user, which is shown 
in Figure \ref{fig:puts:os}. These results show a Zipf-like behavior,
that is, a power law with a small {\em bump} in the lowest
values. After testing the Generalized Extreme Value distribution and
failing for the new batch of experiments, we have fitted it to a
Weibull distribution \cite{thoman1969inferences} with the resulting
parameters shown in Table \ref{tab:puts:os}. 
%
\begin{table}
\caption{Weibull distribution  parameters of the fit of
  the number of {\tt PUT}s per unique IP. \label{tab:puts:os}}
\begin{center}
\begin{tabular}{l|cc}
\hline
Experiment  &  Scale $\sigma$ & Shape $\xi$ \\
\hline
4/4 &  43.07 $\pm$ 5.80 &  0.57 $\pm$ 0.03 \\
4/24 & 22.97 $\pm$ 1.57 & 0.66 $\pm$  0.02  \\
7/31 &  53.18 $\pm$ 7.77 &  0.54 $\pm$ 0.03   \\
\hline
Cache 128 & 205.28 $\pm$ 32.10 & 0.77 $\pm$ 0.07 \\ 
Cache 64 & 178.99 $\pm$ 36.44 & 0.60 $\pm$ 0.05 \\ 
Cache 32 & 168.15 $\pm$ 49.28 & 0.57 $\pm$ 0.07 \\
\hline
% from plot-zipf-openshift.R and ips-puts-alife.R
\end{tabular}
\end{center}
\end{table}
%
The inverse
Weibull distribution is a special case of the GEV distribution
in those papers, and appears usually in natural sciences and
artificial life, usually related to decay. It has been frequently
fitted to volunteer computing frameworks such as SETI@home
\citep{javadi2009mining}. The model that user behavior follows can be
explained straighforwardly: when users visit the page, it draws their
attention for a limited amount of time. They give it a chance for a
few seconds. If something there amuses them or they can engage in a
conversation about it, they stay for a while longer, otherwise, they
leave. The {\em scale} parameter, which is around 20-40 in the first
batch for the 40 Trap problem and between 160 and 200 in the 60 traps
problem, depends mainly on the maximum number of generations people
leave it running. Since it finishes or stalls after a number of
generations shown in Figure \cite{tab:summary:os}, volunteers just
leave after that. Curiously enough, the scale parameter is roughly
twice the median number of {\tt PUT}s per experiment, showing that, on
average, the most loyal users reload the page twice after finishing or
after seeing the evolution does not progress. This rule of thumb
breaks down with the last experiment, however, which is interesting by
itself, too.
